### 2015-10-01 Master Thesis/bibliography.tex
Add Kato to bibliography

\begin{thebibliography}{99999}

\bibitem[Sp96]{Sp96}
	M. Spiess (1996)
	\textit{An arithmetic proof of Pop's Theorem concerning Galois groups of function fields over number fields}.
    J. reine angew. Math. \textbf{478}, 107\---126.

\bibitem[Ka86]{Ka86}
    K. Kato (1986)
    \textit{A Hasse principle for two dimensional global fields, with an appendix by J.-L. Colliot Thélène.} J. reine angew. Math. \textbf{366}, 142-183.

\end{thebibliography}