### 2015-11-21 Seminar Schnitttheorie/article.tex
Add template

\documentclass[10pt,a4paper]{article}
\usepackage{fontspec}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\author{Yichuan Shen}
\title{Vektorbündel \& Gysin-Homomorphismus}
\begin{document}
\maketitle

Hello World!
\end{document}