### 2017-03-13 Master Thesis/bibliography.tex
Add items to bibliography

\begin{thebibliography}{999999}

\bibitem[EP05]{EP05}
    A.J. Engler, A. Prestel (2005)
    \textit{Valued Fields}.
    Springer Monographs in Mathematics.

\bibitem[Sp96]{Sp96}
	M. Spiess (1996)
	\textit{An arithmetic proof of Pop's Theorem concerning Galois groups of function fields over number fields}.
    J. reine angew. Math. \textbf{478}, 107\---126.

\bibitem[Ka86]{Ka86}
    K. Kato (1986)
    \textit{A Hasse principle for two dimensional global fields, with an appendix by J.-L. Colliot Thélène}.
    J. reine angew. Math. \textbf{366}, 142\---183.

\bibitem[ZS60]{ZS60}
    O. Zariski, P. Samuel (1960)
    \textit{Commutative Algebra, Volume II}. 
    Van Nostrand.

\end{thebibliography}