### 2017-05-06 Master Thesis/bibliography.tex
Add introduction

\begin{thebibliography}{999999}

\bibitem[EP05]{EP05}
    A.J. Engler, A. Prestel
    \textit{Valued Fields}.
    Springer Monographs in Mathematics (2005).

\bibitem[Ka86]{Ka86}
    K. Kato
    \textit{A Hasse principle for two dimensional global fields, with an appendix by J.-L. Colliot Thélène}.
    J. reine angew. Math. \textbf{366} (1986), 142\---183.

\bibitem[Ne69]{Ne69}
    J. Neukirch
    \textit{Kennzeichnung der $p$-adischen und der endlichen algebraischen Zahlkörper}.
    Invent. Math. \textbf{6} (1969), 296\---314.

\bibitem[Ne77]{Ne77}
    J. Neukirch
    \textit{Über die absoluten Galoisgruppen algebraischer Zahlkörper}. Asterisque, \textbf{41\---42}, Soc. Math. France, Paris (1977), 67\---79.

\bibitem[Po90]{Po90}
    F. Pop
    \textit{On the Galois theory of function fields of one variable over number fields}.
    J. reine angew. Math. \textbf{406} (1990), 200\---218.

\bibitem[Po94]{Po94}
    F. Pop
    \textit{On Grothendieck's conjecture of birational anabelian geometry}.
    Ann. Math. \textbf{138} (1994), 145\---182.

\bibitem[Po95]{Po95}
    F. Pop
    \textit{On Grothendieck's conjecture of birational anabelian geometry II}, Preprint (1995).

\bibitem[Sp96]{Sp96}
	M. Spiess
	\textit{An arithmetic proof of Pop's Theorem concerning Galois groups of function fields over number fields}.
    J. reine angew. Math. \textbf{478} (1996), 107\---126.

\bibitem[Uc76]{Uc76}
    K. Uchida
    \textit{Isomorphisms of Galois groups}.
    J. Math. Soc. Japan \textbf{28} (1976), no. 4, 617\---620.

\bibitem[Uc77]{Uc77}
    K. Uchida
    \textit{Isomorphisms of Galois groups of algebraic function fields}.
    Ann. Math. (2) \textbf{106} (1977), no. 3, 589\---598.

\bibitem[ZS60]{ZS60}
    O. Zariski, P. Samuel
    \textit{Commutative Algebra, Volume II} (1960). 
    Van Nostrand.

\end{thebibliography}