### 2017-05-10 Master Thesis/thesis.tex
Add abstracts

\documentclass[11pt, a4paper, openany]{memoir}
\usepackage{fontspec}
\usepackage[english]{babel}
\usepackage{amsmath, amsfonts, amssymb, amsthm}
\usepackage{paralist, tikz}
\usetikzlibrary{cd, babel}
\usepackage[inner=4cm, outer=3cm, top=3cm, bottom=3cm]{geometry}
\usepackage{hyperref}
\hypersetup{bookmarksopen=true, pdfstartview=FitH}
\linespread{1.1}
\author{Yichuan Shen}
\title{Galois Characterization of Function Fields of \\ Transcendence Degree $1$ over Number Fields}

\begin{document}

\theoremstyle{plain}
\theoremstyle{definition}
\newtheorem{theorem}{Theorem}[chapter]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem*{definition}{Definition}
\newtheorem*{example}{Example}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{step}{Step}

\frontmatter
\pagenumbering{gobble}

\makeatletter
\begin{center}
\vspace*{0cm}
\begin{large}
Universität Heidelberg\\
Fakultät für Mathematik und Informatik\\
Studiengang Mathematik\\
\vspace{8mm}
\end{large}
\vfill
\begin{large}
\textbf{MASTER THESIS}\\
\end{large}
\vspace{10mm}
\linespread{2}\selectfont
\begin{huge}\@title \end{huge}\\
\linespread{1.1}\selectfont
\vspace{10mm}
by \textsc{yichuan shen}\\
\vspace{3cm}
\vfill
Supervisor: \textsc{prof. dr. alexander schmidt}\\
\today
\end{center}
\makeatother

\clearpage
\
\clearpage

\vspace*{0cm}
\vfill
\paragraph{Declaration.} I declare that I carried out this master thesis independently, and only with the cited sources, literature, and other professional sources. This thesis was not used in the same or in a similar version to achieve an academic grading or is being published elsewhere.\\
\vspace{7mm}\\
\rule{7cm}{0.4pt}
\clearpage

\vspace*{0cm}
\vfill
\renewcommand{\abstractname}{Abstract}
\begin{abstract}
In this thesis we will follow M. Spiess' argument and present his purely arithmetic proof of the fact that isomorphisms between the absolute Galois groups of function fields of transcendence degree $1$ over number fields are induced by field isomorphisms between these fields.
\end{abstract}

\vspace{1cm}

\renewcommand{\abstractname}{Zusammenfassung}
\begin{abstract}
In dieser Arbeit wollen wir M. Spiess' rein arithmetischer Beweis nachgehen und zeigen, wieso Isomorphismen zwischen den absoluten Galoisgruppen von Funktionenkörpern vom Transzendenzgrad $1$ über Zahlkörper schon von Körperisomorphismen zwischen diesen Körpern induziert werden.
\end{abstract}
\vfill
\clearpage

\pagenumbering{arabic}
\setcounter{page}{5}
\tableofcontents

\input{introduction.tex}

\mainmatter
\setcounter{page}{9}

\input{synopsis.tex}
\input{bibliography.tex}
\end{document}